---
title: "Scenith analysis – Alba Perez Arribas"
author: "celbaek"
date: "2025-11-25"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 7,
  fig.height = 5
)

rm(list = ls())

library(flowCore)
library(flowViz)
library(ggcyto)
library(ggridges)
library(openCyto)
library(scales)
library(tidyverse)
library(easystats)
library(ggpubr)
library(ggsignif)
library(viridis)
library(svglite)
library(ggbeeswarm)
library(DT)

# Consistent color scheme across figures
treatment_cols <- c(
  "DMEM"         = "#4D4D4D",
  "Co"         = "#7B1FA2",
  "DG"          = "#D32F2F",
  "2DG4h"        = "#00897B",
  "3PO"       = "#1976D2",
  "O"  = "#9E9E9E",
  "DGO" = "blue"
)

genotype_cols <- c(
  "WT"      = "#1B9E77",
  "WT_HSV"  = "#7570B3",
  "KO"      = "#D95F02",
  "KO_HSV"  = "#E7298A"
)

# Pedagogical base theme
theme_alba <- function(base_size = 12){
  theme_bw(base_size = base_size) +
    theme(
      plot.title       = element_text(face = "bold", size = base_size + 2),
      plot.subtitle    = element_text(size = base_size),
      plot.caption     = element_text(size = base_size - 2, colour = "grey40"),
      strip.background = element_rect(fill = "grey95"),
      strip.text       = element_text(face = "bold"),
      panel.grid.minor = element_blank()
    )
}
```

```{r js_download_helper, echo=FALSE, results='asis'}
cat('
<script>
document.addEventListener("DOMContentLoaded", function() {
  document.querySelectorAll(".download-plot").forEach(function(btn) {
    btn.addEventListener("click", function() {
      var plotContainer = btn.previousElementSibling;
      if (!plotContainer) return;
      var img = plotContainer.querySelector("img");
      if (!img) return;
      var a = document.createElement("a");
      a.href = img.src;
      a.download = btn.getAttribute("data-filename") || "figure.png";
      document.body.appendChild(a);
      a.click();
      document.body.removeChild(a);
    });
  });
});
</script>
')
```

## Experimental overview

This document summarizes a **Scenith assay** performed on cells subjected to metabolic perturbations:

- **DG** (2-deoxy-D-glucose; glycolysis inhibitor)  
- **Oligomycin A** (ATP synthase inhibitor; blocks oxidative phosphorylation)  
- **DGO** (combined DG + oligomycin; “Harringtonine” condition)
- Labelled with **puromycin** (for translation activity) for __xx__ minutes
- Appropriate **DMSO controls** (2 different volumes) and **UNST** (unstained) conditions  

Metabolic activity is read out using **puromycin** (APC channel), which reflects translation and thus ATP availability, and a **live/dead dye** in **FITC**.  

In this report we:

1. Import and annotate FCS files from a 96-well V-bottom plate  
2. Apply a simple gating strategy (singlets (FSC-H and FSC-A side scatter) → live cells (negative Live/Dead) → puromycin-level)  
3. Perform QC on cell numbers and puromycin signal
4. Visualize puromycin distributions across treatments and genotypes  

---

## Data import and sample annotation

We first load all FCS files from the experiment folder and attach plate layout information (genotype and treatment) based on well position.

Glimpse into dataset:

```{r import_fcs}
fcs.dir <- "96 Well - V bottom APA 3/"

df <- read.flowSet(
  path               = fcs.dir,
  pattern            = "\\.fcs$",
  alter.names        = TRUE,
  name.keyword       = "SAMPLE ID",
  truncate_max_range = FALSE
)

#sampleNames(df)
head(pData(phenoData(df)),5)
colnames(df)

df
```

### Sample-to-well mapping

We extract the **well code** from the FCS file names (e.g. `D04` from `Specimen_004_D4_D04.fcs`) and map rows/columns to genotype and treatment.

```{r sample_map}
library(stringr)
library(dplyr)
library(tibble)
library(readxl)
meta <- read_xlsx("Trial3_scenith_apa.xlsx")
meta
sample_map <- tibble(
  sample = sampleNames(df)
) %>%
  mutate(
    well_code = str_extract(sample, "[A-H]\\d{2}"),
    well_row  = str_sub(well_code, 1, 1),
    well_col  = as.integer(str_sub(well_code, 2, 3))
  ) %>% left_join(.,meta)
sample_map



DT::datatable(
  sample_map,
  options = list(pageLength = 10),
  caption = "Mapping of FCS files to plate wells, genotypes and treatments"
)
```

---

## Gating strategy

We apply a sequential gating strategy:

1. **Singlets**: FSC-A vs FSC-H polygon  
2. **Live cells**: low FITC-A (exclude dead cells)  
3. **Puromycin-positive**: APC-A above a fixed threshold  

This section documents the gates and shows QC plots for representative samples.

### Singlet gate (FSC-A vs FSC-H)

```{r gate_singlets}
sqrcut <- matrix(
  c(8000,  20000,
    5000,      0,
   200000,  110000,
   200000,  140000),
  ncol = 2,
  byrow = TRUE
)
colnames(sqrcut) <- c("FSC.A", "FSC.H")

pg <- polygonGate(filterId = "Singlets", gate = sqrcut)
singleCell <- Subset(df, pg)
```

#### QC: Singlet gate in a representative sample

```{r qc_singlets_plot}
p_qc_singlets <- ggcyto(df[1], aes(x = FSC.A, y = FSC.H)) +
  geom_hex(bins = 60) +
  geom_gate(pg, colour = "red", size = 0.6) +
  geom_stats() +
  scale_fill_viridis_c(option = "magma") +
  theme_alba() +
  labs(
    title    = "Step 1: Singlet gate on FSC-A vs FSC-H",
    subtitle = "Red polygon indicates the region used to select single cells",
    x        = "FSC-A",
    y        = "FSC-H",
    fill     = "Cell density"
  )

p_qc_singlets
```

```{r qc_singlets_download, echo=FALSE, results='asis'}
cat('<button class="download-plot" data-filename="qc_singlets_plot.png">
Click here to download the figure
</button>')
```


#### FSC-A vs FSC-H for **all** samples (raw data)

This plot shows the FSC-A vs FSC-H distribution **before any filtering** for all samples, using hexagonal binning.  
Each facet corresponds to one sample and is labelled with sample ID, genotype, treatment and the number of raw events.

```{r fsc_all_samples, fig.height=12, fig.width=18}
library(sp)  # for point.in.polygon

# build combined FSC data for all samples (raw df)
all_fsc <- purrr::map_df(seq_along(df), function(i){
  tibble(
    FSC.A  = exprs(df[[i]])[, "FSC.A"],
    FSC.H  = exprs(df[[i]])[, "FSC.H"],
    sample = sampleNames(df)[i]
  )
}) %>%
  left_join(sample_map, by = "sample")

# function to test if points are inside polygon
inside_gate <- function(df_points, gate_mat) {
  sp::point.in.polygon(
    df_points$FSC.A,
    df_points$FSC.H,
    gate_mat[, 1],
    gate_mat[, 2]
  ) > 0
}

# compute % singlets AND n_raw for each sample using the same df
percent_singlets <- purrr::map_df(seq_along(df), function(i) {
  dat <- as_tibble(exprs(df[[i]])[, c("FSC.A", "FSC.H")])
  inside <- inside_gate(dat, sqrcut)
  tibble(
    sample       = sampleNames(df)[i],
    pct_singlets = 100 * mean(inside),
    n_raw        = nrow(dat)
  )
})

# merge into all_fsc and build the facet label
all_fsc2 <- all_fsc %>%
  left_join(percent_singlets, by = "sample") %>%
  mutate(
    sample_label = paste0(sample, " | ", genotype, " | ", treatment, " | n=", n_raw)
  )

# make a polygon data.frame from the singlet gate (close the polygon)
gate_df <- as.data.frame(sqrcut)
colnames(gate_df) <- c("FSC.A", "FSC.H")
gate_df <- rbind(gate_df, gate_df[1, ])  # close polygon

p_fsc_all <- ggplot(all_fsc2, aes(x = FSC.A, y = FSC.H)) +
  geom_hex(bins = 35) +

  # draw the singlet gate
  geom_polygon(
    data        = gate_df,
    aes(x = FSC.A, y = FSC.H),
    inherit.aes = FALSE,
    color       = "red",
    linewidth   = 0.4,
    fill        = NA
  ) +

  scale_fill_viridis_c(option = "magma") +

  facet_wrap(~ sample_label, scales = "free", ncol = 8) +

  # add percentage label in each facet (top-left corner)
  geom_text(
    data = dplyr::distinct(all_fsc2, sample_label, pct_singlets),
    aes(x = -Inf, y = Inf,
        label = paste0(round(pct_singlets, 1), "% singlets")),
    inherit.aes = FALSE,
    hjust = -0.1, vjust = 1.2,
    size = 2.5,
    color = "red",
    fontface = "bold"
  ) +

  theme_alba(8) +
  labs(
    title    = "FSC-A vs FSC-H for all samples (raw data)",
    subtitle = "Hex-binned FSC-A/FSC-H before gating; red polygon = singlet gate; text shows % singlets per sample",
    x        = "FSC-A",
    y        = "FSC-H",
    fill     = "Cell density"
  )

p_fsc_all


```

```{r fsc_all_samples_download, echo=FALSE, results='asis'}
cat('<button class="download-plot" data-filename="fsc_all_samples.png">
Click here to download the figure
</button>')
```



### Live-cell gate (FITC-A)

We gate out cells with high BV421.A (dead cells) by keeping only events below a fixed threshold.

```{r gate_live}
fitc_threshold <- 20000
rg_fitc <- rectangleGate("BV421.A" = c(0, fitc_threshold))
singleCell_live <- Subset(singleCell, rg_fitc)
```

#### QC: Live/dead distribution in a representative sample

```{r qc_live_plot}

fitc_max <- max(exprs(singleCell[[56]])[, "BV421.A"], na.rm = TRUE)

p_qc_live <- ggcyto(singleCell[56], aes(x = FITC.A)) +
  
  # Blue shading for KEPT cells
  annotate(
    "rect",
    xmin = 1,
    xmax = fitc_threshold,
    ymin = -Inf,
    ymax = Inf,
    fill = "#1976D2",
    alpha = 0.15
  ) +
  
  # Red-striped shading for DISCARDED cells
  annotate(
    "rect",
    xmin = fitc_threshold,
    xmax = fitc_max,
    ymin = -Inf,
    ymax = Inf,
    fill = "red",
    alpha = 0.10
  ) +
  
  # Hatch lines for discarded region
  annotate(
    "segment",
    x = seq(fitc_threshold, fitc_max, length.out = 20),
    xend = seq(fitc_threshold, fitc_max, length.out = 20) + 500,
    y = -Inf,
    yend = Inf,
    colour = "red",
    alpha = 0.15,
    linewidth = 0.5
  ) +
  
  # Density curve
  geom_density(alpha = 0.6, fill = "#1976D2") +
  
  geom_vline(xintercept = fitc_threshold,
             colour = "red",
             linetype = "dashed",
             linewidth = 0.7) +
  
  scale_x_log10() +
  theme_alba() +
  labs(
    title    = "Step 2: Live-cell gate on BV421.A",
    subtitle = "Blue = live cells kept; Red = dead cells discarded (high BV421.A)\nDashed line shows threshold",
    x        = "BV421.A (log10)",
    y        = "Density"
  );p_qc_live

```

```{r qc_live_download, echo=FALSE, results='asis'}
cat('<button class="download-plot" data-filename="qc_live_plot.png">
Click here to download the figure
</button>')
```

---

## QC: cell counts through gating

Here we summarize how many cells are retained at each gating step  
(raw → singlets → live singlets) for every sample.

```{r cell_counts}
samples <- sampleNames(df)

raw_counts      <- purrr::map_int(seq_along(df),              ~ nrow(exprs(df[[.x]])))
singlet_counts  <- purrr::map_int(seq_along(singleCell),      ~ nrow(exprs(singleCell[[.x]])))
live_counts     <- purrr::map_int(seq_along(singleCell_live), ~ nrow(exprs(singleCell_live[[.x]])))

cell_counts <- tibble(
  sample         = samples,
  n_raw          = raw_counts,
  n_singlets     = singlet_counts,
  n_live_singlet = live_counts
) %>%
  left_join(sample_map, by = "sample")
```

### Cell counts after singlet + live gating

```{r cell_counts_plot}
p_cell_counts <- cell_counts %>%
  ggplot(aes(x = reorder(sample, n_live_singlet), y = n_live_singlet, fill = genotype)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = genotype_cols, na.value = "grey80") +
  theme_alba() +
  labs(
    title    = "QC: Number of cells remaining after singlet + live gating",
    subtitle = "Bars are coloured by genotype; samples are ordered by the number of retained cells",
    x        = "Sample",
    y        = "# Live singlet cells",
    fill     = "Genotype"
  )

p_cell_counts
```

```{r cell_counts_download, echo=FALSE, results='asis'}
cat('<button class="download-plot" data-filename="cell_counts_plot.png">
Click here to download the figure
</button>')
```

### Interactive table of cell counts

```{r cell_counts_table}
DT::datatable(
  cell_counts %>%
  mutate(genotype = factor(genotype,levels =c("WT","WT_HSV","KO","KO_HSV"))) %>% 
    arrange(genotype, treatment, sample),
  options = list(pageLength = 10),
  caption = "Cell counts at each gating step per sample"
)
```


## Puromycin (APC-A)

```{r gate_puro}
puro_threshold <- 1000
gate_puro <- rectangleGate("APC.A" = c(puro_threshold, Inf))
puro_cells <- Subset(singleCell_live, gate_puro)
```

We take a look at translation (threshold of >80 - can be adjusted)


#### QC: Puromycin vs FSC in representative samples

```{r qc_puro_plot}
plot_list <- c(2,3,4,5,6,7,8)
plot_list <- c(69,73,74,21,25,26)

p_qc_puro <- ggcyto(singleCell_live[plot_list], aes(x = FSC.A, y = APC.A)) +
  geom_hex(bins = 60) +
  scale_y_log10(limits = c(1, NA)) +
  geom_hline(yintercept = puro_threshold, color = "red", linetype = "dashed") +
  scale_fill_viridis_c(option = "magma") +
  theme_alba() +
  facet_wrap(~ name) +
  labs(
    title    = "Step 3: Puromycin gate on APC-A",
    subtitle = "Dashed line marks the puromycin-positive threshold (APC-A > 1000)",
    x        = "FSC-A",
    y        = "APC-A (log10)",
    fill     = "Cell density"
  )

p_qc_puro
```

```{r qc_puro_download, echo=FALSE, results='asis'}
cat('<button class="download-plot" data-filename="qc_puro_plot.png">
Click here to download the figure
</button>')
```





---

## Puromycin summary per sample

We now summarize, per sample:

- Number of live singlet cells  
- Number and percentage of puromycin-positive cells  
- Mean APC-A in all live cells and in puromycin-positive cells  

```{r puro_summary}
sample_ids <- sampleNames(singleCell_live)

puro_summary <- purrr::map_df(seq_along(sample_ids), function(i) {
  ff_live <- singleCell_live[[i]]
  ff_puro <- puro_cells[[i]]

  tibble(
    sample         = sample_ids[i],
    n_live         = nrow(exprs(ff_live)),
    n_puro         = nrow(exprs(ff_puro)),
    pct_puro       = 100 * nrow(exprs(ff_puro)) / nrow(exprs(ff_live)),
    mean_puro_live = mean(exprs(ff_live)[, "APC.A"], na.rm = TRUE),
    mean_puro_pos  = mean(exprs(ff_puro)[, "APC.A"], na.rm = TRUE)
  )
}) %>%
  left_join(sample_map, by = "sample")
```

### Mean puromycin signal per sample

```{r mean_puro_plots, fig.height=8}
p_mean_live <- ggplot(puro_summary, aes(x = reorder(sample, mean_puro_live),
                                        y = mean_puro_live,
                                        fill = treatment)) +
  geom_col() +
  geom_hline(yintercept = 1000, color = "red", linetype = "dashed") +
  coord_flip() +
  scale_fill_manual(values = treatment_cols, na.value = "grey80") +
  theme_alba() +
  labs(
    title    = "Mean puromycin signal per sample (all live cells)",
    subtitle = "Dashed line indicates the puromycin-positive threshold (APC-A = 1000)",
    x        = "Sample",
    y        = "Mean APC-A",
    fill     = "Treatment"
  )

p_mean_pos <- ggplot(puro_summary, aes(x = reorder(sample, mean_puro_pos),
                                       y = mean_puro_pos,
                                       fill = treatment)) +
  geom_col() +
  geom_hline(yintercept = 1000, color = "red", linetype = "dashed") +
  coord_flip() +
  scale_fill_manual(values = treatment_cols, na.value = "grey80") +
  theme_alba() +
  labs(
    title    = "Mean puromycin signal per sample (puromycin-positive cells)",
    subtitle = "Only cells above the APC-A threshold are included in the mean",
    x        = "Sample",
    y        = "Mean APC-A (puro+)",
    fill     = "Treatment"
  )

p_mean_both <- ggpubr::ggarrange(p_mean_live, p_mean_pos, ncol = 1)

p_mean_both
```

```{r mean_puro_download, echo=FALSE, results='asis'}
cat('<button class="download-plot" data-filename="mean_puro_plots.png">
Click here to download the figure
</button>')
```

### Interactive table of puromycin summary

```{r puro_summary_table}
DT::datatable(
  puro_summary %>%
  mutate(genotype = factor(genotype,levels =c("WT","WT_HSV","KO","KO_HSV"))) %>% 
    arrange(genotype, treatment, sample),
  options = list(pageLength = 10),
  caption = "Puromycin summary statistics per sample"
)
```

---

## Puromycin distributions at the single-cell level

We now go down to **single-cell** resolution to look at APC-A distributions across treatments and genotypes.

### Build cell-level data frame

```{r cell_level_table}
geomfi <- function(x) {
  x <- x[x > 0]
  exp(mean(log(x), na.rm = TRUE))
}

sample_ids_live <- sampleNames(singleCell_live)

cell_level <- purrr::map_df(seq_along(sample_ids_live), function(i) {
  ff <- singleCell_live[[i]]

  as_tibble(exprs(ff)) %>%
    mutate(sample = sample_ids_live[i])
}) %>%
  left_join(sample_map, by = "sample")

# keep only cells with non-trivial APC.A for distribution plots
cell_level_filtered <- cell_level %>%
  mutate(genotype = factor(genotype,levels =c("WT","WT_HSV","KO","KO_HSV"))) %>% 
  filter(APC.A >= puro_threshold) 

# geometric means per genotype × treatment
geo_means <- cell_level_filtered %>%
  mutate(genotype = factor(genotype,levels =c("WT","WT_HSV","KO","KO_HSV"))) %>% 
  filter(treatment %in% c("Co", "DG", "DGO", "O")) %>%
  group_by(genotype, treatment) %>%
  summarise(
    geo_mean = geomfi(APC.A),
    .groups  = "drop"
  )
```



### Scenith-derived metabolic parameters (per genotype)

Here we summarise classical Scenith metrics, computed **per genotype** using the  
geometric mean APC-A in each condition:

- **Glucose dependence** = 100 · (Co − DG) / (Co − DGO)  
- **FAO/AAO capacity**  = 100 − glucose dependence  
- **Mitochondrial dependence** = 100 · (Co − O) / (Co − DGO)  
- **Glycolytic capacity**      = 100 − mitochondrial dependence  

```{r scenith_params_tables}
library(tidyr)

# Parameters based on Co, DG, DGO
scenith_dg <- geo_means %>%
  filter(treatment %in% c("Co", "DG", "DGO")) %>%
  select(genotype, treatment, geo_mean) %>%
  pivot_wider(names_from = treatment, values_from = geo_mean) %>%
  mutate(
    glucose_dependence = 100 * ((Co - DG) / (Co - DGO)),
    fao_aao_capacity   = 100 - glucose_dependence
  )

# Parameters based on Co, O, DGO
scenith_o <- geo_means %>%
  filter(treatment %in% c("Co", "O", "DGO")) %>%
  select(genotype, treatment, geo_mean) %>%
  pivot_wider(names_from = treatment, values_from = geo_mean) %>%
  mutate(
    mito_dependence      = 100 * ((Co - O) / (Co - DGO)),
    glycolytic_capacity  = 100 - mito_dependence
  )

scenith_summary <- scenith_dg %>%
  full_join(scenith_o, by = "genotype", suffix = c("_dg", "_o")) %>%
  select(
    genotype,
    Co_dg = Co_dg,
    DG,
    DGO_dg = DGO_dg,
    glucose_dependence,
    fao_aao_capacity,
    Co_o = Co_o,
    O,
    DGO_o = DGO_o,
    mito_dependence,
    glycolytic_capacity
  )

DT::datatable(
  scenith_summary,
  options = list(pageLength = 10),
  caption = "Scenith-derived parameters per genotype (based on geometric mean APC-A)"
)
```

#### Glucose dependence and FAO/AAO capacity (Co vs DG vs DGO)

For each genotype, we pool technical replicates and show a single density per treatment.  
Arrows indicate the distance between geometric means that define:

- **Glucose dependence** (DG → Co)  
- **FAO/AAO capacity** (DGO → DG)  

```{r dist_co_dg_dgo_params,fig.height=8,fig.width=12}
p_co_dg_dgo_params <- cell_level_filtered %>%
  filter(treatment %in% c("Co", "DG", "DGO")) %>%
  ggplot(aes(x = APC.A, fill = treatment)) +
  
  # Density uses fill only (so fill legend remains)
  geom_density(alpha = 0.6, aes(color=treatment)) +
  
  scale_x_log10(limits=c(1,10e6)) +
  facet_wrap(~ genotype, ncol = 2) +
  
  # Show only fill legend
  scale_fill_manual(values = treatment_cols) +
  
  # Remove color legend entirely (but keep color mapping for vlines)
  scale_colour_manual(values = treatment_cols, guide = "none") +
  
  # geometric mean lines
  geom_vline(
    data        = geo_means %>% filter(treatment %in% c("Co", "DG", "DGO")),
    aes(xintercept = geo_mean, colour = treatment),
    linetype    = "dashed",
    linewidth   = 0.6,
    inherit.aes = FALSE,
    show.legend = FALSE
  ) +
  
  # arrows & labels: glucose dependence (DG -> Co)
  geom_segment(
    data        = scenith_dg,
    aes(x = DG, xend = Co, y = 0.5, yend = 0.5),
    arrow       = arrow(length = unit(0.15, "cm")),
    inherit.aes = FALSE
  ) +
  geom_label(
    data        = scenith_dg,
    aes(x = (DG + Co) / .8, y = 0.2,
        label = paste0("1. Glc dep = ", round(glucose_dependence, 1), "%")),
    size        = 4,
    inherit.aes = FALSE
  ) +
  
  # arrows & labels: FAO/AAO capacity (DGO -> DG)
  geom_segment(
    data        = scenith_dg,color="purple4",
    aes(x = DGO, xend = DG, y = 1, yend = 1),
    arrow       = arrow(length = unit(0.15, "cm")),
    inherit.aes = FALSE
  ) +
  geom_label(
    data        = scenith_dg,color="purple4",
    aes(x = (DGO + DG) / 8, y = 0.8,
        label = paste0("4. FAO/AAO cap = ", round(fao_aao_capacity, 1), "%")),
    size        = 4,
    inherit.aes = FALSE
  ) +
  
  theme_alba(14) +
  labs(
    title    = "Scenith parameters – Glucose dependence and FAO/AAO capacity",
    subtitle = "Densities per treatment pooled across technical replicates; arrows show distances between geometric means",
    x        = "APC-A (log10)",
    y        = "Density",
    fill     = "Treatment"
  )

p_co_dg_dgo_params


```

```{r dist_co_dg_dgo_params_download, echo=FALSE, results='asis'}
cat('<button class="download-plot" data-filename="dist_co_dg_dgo_params.png">
Click here to download the figure
</button>')
```

#### Mitochondrial dependence and glycolytic capacity (Co vs O vs DGO)

Similarly, we pool replicates and visualise:

- **Mitochondrial dependence** (O → Co)  
- **Glycolytic capacity** (DGO → O)  

```{r dist_co_o_dgo_params,fig.width=12,fig.height=8}
p_co_o_dgo_params <- cell_level_filtered %>%
  filter(treatment %in% c("Co", "O", "DGO")) %>%
  ggplot(aes(x = APC.A, fill = treatment)) +
  
  # Use fill for legend, no color mapping here
  geom_density(alpha = 0.6, aes(color=treatment)) +
  
  scale_x_log10(limits=c(1,10e6)) +
  facet_wrap(~ genotype, ncol = 2) +
  
  # Fill legend (Treatment)
  scale_fill_manual(values = treatment_cols) +
  
  # Colour only for vlines, no legend
  scale_colour_manual(values = treatment_cols, guide = "none") +
  
  # geometric mean lines (coloured by treatment, but legend suppressed)
  geom_vline(
    data        = geo_means %>% filter(treatment %in% c("Co", "O", "DGO")),
    aes(xintercept = geo_mean, colour = treatment),
    linetype    = "dashed",
    linewidth   = 0.6,
    inherit.aes = FALSE,
    show.legend = FALSE
  ) +
  
  # arrows & labels: mitochondrial dependence (O -> Co)
  geom_segment(
    data        = scenith_o,
    aes(x = O, xend = Co, y = 0.5, yend = 0.5),
    arrow       = arrow(length = unit(0.15, "cm")),
    inherit.aes = FALSE,
    color       = "blue"
  ) +
  geom_label(
    data        = scenith_o,
    aes(x = (O + Co) / .8, y = 0.2,
        label = paste0("2. Mito dep = ", round(mito_dependence, 1), "%")),
    size        = 4,
    inherit.aes = FALSE,
    color       = "blue"
  ) +
  
  # arrows & labels: glycolytic capacity (DGO -> O)
  geom_segment(
    data        = scenith_o,
    aes(x = DGO, xend = O, y = 1, yend = 1),
    arrow       = arrow(length = unit(0.15, "cm")),
    inherit.aes = FALSE
  ) +
  geom_label(
    data        = scenith_o,
    aes(x = (DGO + O) / 5, y = 0.8,
        label = paste0("3. Glyc cap = ", round(glycolytic_capacity, 1), "%")),
    size        = 4,
    inherit.aes = FALSE
  ) +
  
  theme_alba(14) +
  labs(
    title    = "Scenith parameters – Mitochondrial dependence and glycolytic capacity",
    subtitle = "Densities per treatment pooled across technical replicates; arrows show distances between geometric means",
    x        = "APC-A (log10)",
    y        = "Density",
    fill     = "Treatment"
  )

p_co_o_dgo_params

```

```{r dist_co_o_dgo_params_download, echo=FALSE, results='asis'}
cat('<button class="download-plot" data-filename="dist_co_o_dgo_params.png">
Click here to download the figure
</button>')
```

#### Bar plots
```{r puro_bar_genotype,fig.width=12,fig.height=6}
# Prepare data: geometric mean APC-A per genotype × treatment
bar_puro <- geo_means %>%
  filter(treatment %in% c("Co", "DG", "O", "DGO")) %>%
  mutate(
    treatment = factor(treatment, levels = c("Co", "DG", "O", "DGO"))
  )

p_puro_bar <- ggplot(bar_puro,
                     aes(x = treatment,
                         y = geo_mean,
                         fill = treatment)) +
  geom_col(width = 0.7) +
  facet_wrap(~ genotype, ncol = 4) +
  scale_fill_manual(values = c(
    "Co"  = "#4D4D4D",  # grey
    "DG"  = "#7B1FA2",  # purple
    "O"   = "#D32F2F",  # red
    "DGO" = "#00897B"   # green from earlier
  )) +
  theme_alba(14) +
  labs(
    title    = "Translation per condition and genotype",
    subtitle = "Bars show geometric mean puromycin signal (APC-A) in live singlet cells; technical replicates pooled",
    x        = "Condition",
    y        = "Geometric mean APC-A (puromycin MFI)",
    fill     = "Treatment"
  )

p_puro_bar

```


---

## Notes for interpretation

- **DG** is expected to reduce ATP production via glycolysis, leading to reduced puromycin incorporation.  
- **Oligomycin** blocks mitochondrial ATP, with a complementary impact.  
- **DGO (Harringtonine)** combines both and should show the strongest reduction in puromycin signal.  
- Differences between **WT / WT_HSV / KO / KO_HSV** genotypes can be read out as shifts in the puromycin distributions and the derived Scenith parameters.  

Each figure above is embedded in the HTML and can be downloaded as a PNG directly from the report using the buttons.
